---
title: "Summary of Food Prices for Turkey"
author: "Data R Sizlari"
date: "Dec 2, 2018"
output: html_document
---

```{r, echo=FALSE, message=FALSE}
library(tidyverse)
library(ggplot2)
library(forcats)
library(scales)
library(lubridate)
```

## Download Raw Data

Our raw excel file is in our project repository. We can automatically download that file and put it in a temporary file. Then we can read that excel document into R and remove the temp file.

```{r}
# Create a temporary file
tmp=tempfile(fileext=".xlsx")

# Download file from repository to the temp file
download.file("https://github.com/MEF-BDA503/gpj18-data-r-sizlari/blob/master/wfp_food_prices_turkey.csv?raw=true",destfile=tmp,mode='wb')

# Read that csv file.
raw_data=read.csv(tmp,skip=1)

# Remove the temp file
file.remove(tmp)
```

## Data Preprocessing

```{r}
# Remove unwanted columns, adjust column names and delete the rows type = "non-food" 
raw_data=
raw_data %>%
select(-X.country.name,-X.adm1.name,-X.adm1.code,-X.item.code,-X,-X.1,-X.2,-X.item.type.code,-X.meta.id,-X.3)
colnames(raw_data) = c("date","name","unit","type","value","currency","market")
raw_data = subset(raw_data,subset=type != "non-food")
```

```{r}
# Add month and year columns 
raw_data=
raw_data %>%
mutate(date = ymd(date)) %>% 
mutate_at(vars(date), funs(year, month)) 

# Save organized file as ".rds"
saveRDS(raw_data, file = "C:/Ufuk/BDA_503_Data_Analytics_Essentials/Project/Dataset/wfp_food_prices_turkey.rds")

# Reading organized ".rds" file from GPJ 
# Create a temporary file
tmp=tempfile(fileext=".rds")
# Download file from repository to the temp file
download.file("https://github.com/MEF-BDA503/gpj18-data-r-sizlari/blob/master/wfp_food_prices_turkey.rds?raw=true",destfile=tmp,mode='wb')
# Read that rds file.
raw_data=read_rds(tmp)
# Remove the temp file
file.remove(tmp)
head(raw_data)
```

## Data Summary after Preprocessing

This dataset contains Food Prices data for Turkey. Food prices data comes from the World Food Programme (WFP) and covers 52 different kinds of foods within 8 main categories such as cereals and tubers, milk and dairy, oil and fats and vegetables and fruits for Turkey grouped as national average, Ankara, Istanbul and Izmir. It is updated weekly but contains to a large extent monthly data. The data goes back to May 15, 2013 and includes 4,718 rows and 15 columns. The main numerical information in dataset is the food price. The data is basically categorized into food category, market, months and units.

After organizing the data, the details of the column names & descriptions are represented below:

*	date: Date
*	name : Food name
*	unit : Unit of measure
*	type : Food category
*	value : Food Price
*	currency : Curreny (TRY)
*	market : Related market name (Ankara, Istanbul, Izmir and National Average)
*	year : Year
* month : Month 

## Yearly Average Price per Food Type

```{r}
    raw_data %>% 
    group_by(type, year) %>%
    summarise(yearly_average_price=mean(value)) %>%
    ggplot(data = ., aes(x = type, y = yearly_average_price, 
    fill = as.character(year))) + geom_bar(stat = "identity") + aes(x = reorder(type, -yearly_average_price), 
    y =   yearly_average_price) + labs(x = "", y = "", title = "Yearly Average Price per Food Type") + theme_bw() + theme( axis.text.x = element_text(angle = 45, 
    vjust = 0.5, hjust = 0.5, size = 12)) + scale_y_continuous(labels = scales::comma) +   guides(fill=guide_legend(title="Year", reverse=TRUE)) 
```

## Overall Average Price based on Market

```{r}
    raw_data %>% 
    group_by(type, market) %>%
    summarise(market_average_price=mean(value)) %>%
    ggplot(data = ., aes(x = type, y = market_average_price, 
    fill = as.character(market))) + geom_bar(stat = "identity") + aes(x = reorder(type, -market_average_price), 
    y =   market_average_price) + labs(x = "", y = "", title = "Overall Average Price based on Market") + theme_bw() + theme( axis.text.x = element_text(angle = 45, 
    vjust = 0.5, hjust = 0.5, size = 12)) + scale_y_continuous(labels = scales::comma) +     guides(fill=guide_legend(title="Market", reverse=TRUE)) 
```

## Average Price Development of Milk and Dairy

```{r}
    raw_data %>% 
    filter(type=="milk and dairy") %>%
    group_by(name, year) %>%
    summarize(average_price=mean(value)) %>%
    ggplot(data = ., aes(x = name, y = average_price, 
    fill = as.character(year))) + geom_bar(stat = "identity") + aes(x = reorder(name, -average_price), 
    y =   average_price) + labs(x = "", y = "", title = "Average Price Development of Milk and Dairy") + theme_bw() + theme( axis.text.x = element_text(angle = 45, 
    vjust = 0.5, hjust = 0.5, size = 12)) + scale_y_continuous(labels = scales::comma) +   guides(fill=guide_legend(title="Year", reverse=TRUE)) 

```

## The products with maximum price change

```{r}
price_change = raw_data %>%
  filter(market=="National Average")%>%
  group_by(name, year) %>%
  summarise(yearly_average_price=mean(value)) %>%
  mutate(YoY_growth = (yearly_average_price/lag(yearly_average_price))^(1/(year-lag(year))) - 1) %>%
  arrange(desc(YoY_growth))
price_change
```


```{r}
price_change %>% 
filter(name=="Eggplants - Retail" | name=="Onions - Retail" | name=="Cucumbers (greenhouse) - Retail" | name=="Potatoes - Retail" | name== "Oil (olive) - Retail" ) %>% 
  ggplot(aes(x = year, y = yearly_average_price, group=name, colour=name  )) + 
  geom_line() + 
  geom_point()+
  labs( x="Year" , y = "Yearly Average Price")+ 
  theme(legend.position = "bottom", axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 0.0, size = 10)) 

```


## References:

*	[Food Prices for Turkey Dataset](https://data.humdata.org/dataset/wfp-food-prices-for-turkey/resource/33f303b8-4557-424e-8205-227b889f00e5)
*	[WFP](https://data.humdata.org/organization/3ecac442-7fed-448d-8f78-b385ef6f84e7)










